---
title: "Music Popularity K Means Algorithm"
author: "Rod Rosales"
date: "12/5/2021"
output: html_document
---

#Library

```{r}
library(cluster)
library(caret)
library(factoextra)
library(plyr)
library(dplyr)
library(rgl)
library(plot3D)
library(plotly)
```


###### K-Means Cluster Analysis (Music Popularity)########

#1) Loading Music Genre Data
```{r}
Music_Gen <- read.csv("C:/Users/rod_r/OneDrive/Desktop/Syracuse Classes/Fall 2021/IST 707 Data Analytics/Project/music_genre.csv")
```

#2) Rename Columns
```{r}
colnames(Music_Gen) <- c("ID", "Artist", "Song", "Popularity", "Acoustics", "Danceability", "Duration_ms","Energy", "Instrumentalness", "Key", "Liveness", "Loudness", "Mode", "Lyrics", "Tempo","Obtained_Date", "Valence", "Genre")
```

#3) Get Rid of ID, Artist, and unwanted variables
```{r}
Music_Gen <- Music_Gen[,-1:-3]
Music_Gen$Obtained_Date <- NULL
Music_Gen$Duration_ms <- NULL
```

#4) Combine Rap and Hip-Hop Into One Music Genre
```{r}
Music_Gen$Genre <- gsub("Rap","Hip-Hop",Music_Gen$Genre)
Music_Gen$Genre <- gsub("Hip-Hop","Rap/Hip-Hop",Music_Gen$Genre)
table(Music_Gen$Genre)
```

#5) Remove Entries with values we can't work with and numerate non numerical variables
```{r}
Music_Gen <-
Music_Gen %>%
  mutate_if(is.numeric, ~replace(., . == -1, NA))

Music_Gen <-
Music_Gen %>%
  mutate_if(is.numeric, ~replace(., . == 0, NA))

Music_Gen <-
Music_Gen %>%
  mutate_if(is.factor, ~replace(., . == "empty_field", NA))
```

#6) Numerate non-Numeric Values
```{r}
Music_Gen$Tempo <- as.numeric(Music_Gen$Tempo)
Music_Gen <- transform(Music_Gen, Key = as.numeric(factor(Key)))
Music_Gen <- transform(Music_Gen, Mode = as.numeric(factor(Mode)))
```

#7) Remove Entries with NA Values
```{r}
Music_Gen <- na.omit(Music_Gen)
```

#8) Get a distribution of the data by Genre
```{r}
Genre.Labels = Music_Gen$Genre
table(Genre.Labels)
```

#9) Remove Genre
```{r}
Music_Gen_Data <- Music_Gen[1:12]
```

#10) Scale the Data
```{r}
Music_Gen_Data_Scale <- scale(Music_Gen_Data)
```

#11) WSS Plot to determine # of clusters
```{r}
wssplot <- function(data, nc=10, seed=1234)
{
  wss <- (nrow(data)-1)*sum(apply(data,2,var))
  for (i in 2:nc){
  set.seed(seed)
  wss[i] <- sum(kmeans(data,centers=i)$withinss)}
  plot(1:nc, wss, type="b", xlab= "Number of Clusters",
  ylab = "Within groups Sum of Squares")
}
wssplot(Music_Gen_Data_Scale)
```

#12) Visualize and label rows
```{r}
km.out <- kmeans(Music_Gen_Data_Scale, centers = 4, nstart = 100)
print(km.out)
```

#13-1) Label Rows and Visualize Cluster Plot
```{r}
km.clusters <- km.out$cluster
rownames(Music_Gen_Data_Scale) <- paste(Music_Gen$Genre, 1:dim(Music_Gen)[1], sep = "_")
Music_Cluster_Plot <- fviz_cluster(list(data = Music_Gen_Data_Scale, cluster = km.clusters))
Music_Cluster_Plot
```

#13-2) Table of Cluster Plot Above
```{r}
table(km.clusters,Music_Gen$Genre)
```

#13-3)  above cluster center (means) table with variables within each cluster
```{r}
km.out$centers
```

#13-4) Individual Variables by Song including cluster assignment
```{r}
cluster_assignment <- data.frame(Music_Gen_Data_Scale,km.clusters)
cluster_assignment
```



###Variables Number Code for 3d Plot Below###
#Popularity : 1
#Acoustics : 2
#Danceability : 3
#Energy : 4
#Instrumentalness : 5
#Key : 6
#Liveness : 7
#Loudness : 8
#Mode : 9
#Lyrics : 10
#Tempo : 11
#Valence : 12
###Variables Number Code for 3d Plot Below###



#14) 3d Plot for Music Popularity
```{r}
plot3d(Music_Gen_Data_Scale[, c(1, 2, 10)], col=rainbow(4)[km.clusters])
legend3d("topright", legend = paste('Cluster', c('1', '2', '3','4')), pch = 16, col = rainbow(4), cex=1, inset=c(0.02))
```






################################## K-Means Cluster     END    ##############################################

