---
title: "Rod Preprocessing + Models"
author: "Rod Rosales"
date: "11/10/2021"
output: html_document
---

#Libraries
```{r}
library(cluster)
library(caret)
library(rpart)
library(rattle)
library(factoextra)
library(dendextend)
library(tm)
library(gridExtra)
library(stringr)
library(tidyr)
library(rpart.plot)
library(C50)
library(textclean)
library("naniar")
library(plyr)
library(dplyr)
library(arules)
library(arulesViz)
library(caTools)
library(stats)
library(ggfortify)
```

#### Pre-Processing ####

#1) Loading Music Genre Data
```{r}
Music_Gen <- read.csv("C:/Users/rod_r/OneDrive/Desktop/Syracuse Classes/Fall 2021/IST 707 Data Analytics/Project/music_genre.csv", encoding='UTF-8')
```

#2) Rename Columns
```{r}
colnames(Music_Gen) <- c("ID", "Artist", "Song", "Popularity", "Acoustics", "Danceability", "Duration_ms","Energy", "Instrumentalness", "Key", "Liveness", "Loudness", "Mode", "Lyrics", "Tempo","Obtained_Date", "Valence", "Genre")
```

#3) Remove Entries with values we can't work with
```{r Music_Gen, echo=TRUE, eval=TRUE}
Music_Gen <-
  Music_Gen %>%
  mutate_if(is.numeric, ~replace(., . == -1, NA))
Music_Gen <-
  Music_Gen %>%
  mutate_if(is.numeric, ~replace(., . == 0, NA))
Music_Gen <-
  Music_Gen %>%
  mutate_if(is.factor, ~replace(., . == "empty_field", NA))
```

#4) Recharacterize Character Variables for statistics
```{r}
Music_Gen$ID <- as.factor(Music_Gen$ID)
Music_Gen$Artist <- as.factor(Music_Gen$Artist)
Music_Gen$Song <- as.factor(Music_Gen$Song)
Music_Gen$Key <- as.factor(Music_Gen$Key)
Music_Gen$Mode <- as.factor(Music_Gen$Mode)
Music_Gen$Genre <- as.factor(Music_Gen$Genre)
Music_Gen$Tempo <- as.numeric(Music_Gen$Tempo)
```

#5) Remove Entries with NA Values
```{r}
Music_Gen <- na.omit(Music_Gen)
```

#6) Index the ID numbers
```{r}
rownames(Music_Gen)<-Music_Gen$ID
```

#7) Get Rid of Id, Artist, and non numerical values
```{r}
New_Music_Gen <- Music_Gen[,-1:-3]
New_Music_Gen$Key <- NULL
New_Music_Gen$Mode <- NULL
New_Music_Gen$Genre <- NULL
Music_Gen$Obtained_Date <- NULL
```

#### Pre-Processing END ####



#############################Aproproi Algorithm#############################

#1) Loading Music Genre Data
```{r}
Music_Gen <- read.csv("C:/Users/rod_r/OneDrive/Desktop/Syracuse Classes/Fall 2021/IST 707 Data Analytics/Project/music_genre.csv", encoding='UTF-8')
```

```{r}
View(Music_Gen)
str(Music_Gen)
```

#2) Rename Columns
```{r}
colnames(Music_Gen) <- c("ID", "Artist", "Song", "Popularity", "Acoustics", "Danceability", "Duration_ms","Energy", "Instrumentalness", "Key", "Liveness", "Loudness", "Mode", "Lyrics", "Tempo","Obtained_Date", "Valence", "Genre")
```

#3) Remove Entries with values we can't work with
```{r}
Music_Gen <-
  Music_Gen %>%
  mutate_if(is.numeric, ~replace(., . == -1, NA))
Music_Gen <-
  Music_Gen %>%
  mutate_if(is.numeric, ~replace(., . == 0, NA))
Music_Gen <-
  Music_Gen %>%
  mutate_if(is.factor, ~replace(., . == "empty_field", NA))
Music_Gen$Tempo <- as.numeric(Music_Gen$Tempo)
```

#4) Remove Entries with NA Values
```{r}
Music_Gen <- na.omit(Music_Gen)
```

#5) Index the ID numbers
```{r}
rownames(Music_Gen)<-Music_Gen$ID
```

#6) Get Rid of ID, Artist, and non numerical values
```{r}
New_Music_Gen <- Music_Gen[,-1:-3]
New_Music_Gen$Mode <- NULL
New_Music_Gen$Acoustics <- NULL
New_Music_Gen$Duration_ms <- NULL
New_Music_Gen$Loudness <- NULL
New_Music_Gen$Lyrics <- NULL
New_Music_Gen$Liveness <- NULL
New_Music_Gen$Obtained_Date <- NULL
```

```{r}
View(Music_Gen)
str(Music_Gen)
```

#7) Range of Danceability
```{r}
range(Music_Gen$Danceability)
```

#8) Range of Popularity
```{r}
range(Music_Gen$Popularity)
```

#9) Range of Energy
```{r}
range(Music_Gen$Energy)
```

#10) Range of Tempo
```{r}
range(Music_Gen$Tempo)
```

#11) Range of Instrumentalness
```{r}
range(Music_Gen$Instrumentalness)
```

#13) Range of Valence
```{r}
range(Music_Gen$Valence)
```




#14) Discretize Variables
```{r}
New_Music_Gen$Danceability <- cut(New_Music_Gen$Danceability, breaks = c(.01,.25,.50,.75,Inf),labels=c("Not Danceable","Head-Bobber"," Danceable","Highly Danceable"))
New_Music_Gen$Popularity <- cut(New_Music_Gen$Popularity, breaks = c(1,30,50,70,90,Inf),labels=c("Niche","Not Popular","Somwhat Popular","Popular","Mainstream"))
New_Music_Gen$Energy <- cut(New_Music_Gen$Energy, breaks = c(.0005,.30,.60,.90,Inf),labels=c("Low","Moderate", "Upbeat", "Energetic"))
New_Music_Gen$Tempo <- cut(New_Music_Gen$Tempo, breaks = c(30,50,70,90,150,Inf),labels=c("Grave","Adagietto", "Moderato", "Allegro", "Prestissimo"))
New_Music_Gen$Instrumentalness <- cut(New_Music_Gen$Instrumentalness, breaks = c(.0,.10,.30,.60,.90,Inf),labels=c("Acapella","Low Instruments", "Moderate", "Instruments", "Full Instrumental"))
New_Music_Gen$Valence <- cut(New_Music_Gen$Valence, breaks = c(.01,.30,.50,.70,.90,Inf),labels=c("Depressing","Angry","Moderate","cheerful","Euphoric"))
```

```{r}
View(New_Music_Gen)
str(New_Music_Gen)
```

#15)Recode:
```{r}
New_Music_Gen$Popularity=dplyr::recode(New_Music_Gen$Popularity, YES="married=YES", NO="married=NO")
New_Music_Gen$Danceability=dplyr::recode(New_Music_Gen$Danceability, YES="car=YES", NO="car=NO")
New_Music_Gen$Energy=dplyr::recode(New_Music_Gen$Energy, YES="save_act=YES", NO="save_act=NO")
New_Music_Gen$Instrumentalness=dplyr::recode(New_Music_Gen$Instrumentalness, YES="current_act=YES", NO="current_act=NO")
New_Music_Gen$Tempo=dplyr::recode(New_Music_Gen$Tempo, YES="mortgage=YES", NO="mortgage=NO")
New_Music_Gen$Valence=dplyr::recode(New_Music_Gen$Valence, YES="pep=YES", NO="pep=NO")
New_Music_Gen$Genre=dplyr::recode(New_Music_Gen$Genre, YES="pep=YES", NO="pep=NO")
New_Music_Gen$Key=dplyr::recode(New_Music_Gen$Key, YES="pep=YES", NO="pep=NO")
```

#16) Load transformed data into apriori algorithm
```{r}
MyRules = apriori(New_Music_Gen, parameter = list(supp = 0.001, conf = 0.9, maxlen = 4))
```

#17) Sort and inspect(myRules)
```{r}
MyRules.sorted <- sort(MyRules, by="lift")
inspect(MyRules.sorted)
```

#############################Aproproi Algorithm END#############################











###### K-Means Cluster Analysis (Music Popularity)########

#1) Loading Music Genre Data
```{r}
Music_Gen <- read.csv("C:/Users/rod_r/OneDrive/Desktop/Syracuse Classes/Fall 2021/IST 707 Data Analytics/Project/music_genre.csv", encoding='UTF-8')
```

#2) Rename Columns
```{r}
colnames(Music_Gen) <- c("ID", "Artist", "Song", "Popularity", "Acoustics", "Danceability", "Duration_ms","Energy", "Instrumentalness", "Key", "Liveness", "Loudness", "Mode", "Lyrics", "Tempo","Obtained_Date", "Valence", "Genre")
```

#3) Remove Entries with values we can't work with
```{r}
Music_Gen <-
Music_Gen %>%
  mutate_if(is.numeric, ~replace(., . == -1, NA))

Music_Gen <-
Music_Gen %>%
  mutate_if(is.numeric, ~replace(., . == 0, NA))

Music_Gen <-
Music_Gen %>%
  mutate_if(is.factor, ~replace(., . == "empty_field", NA))
Music_Gen$Tempo <- as.numeric(Music_Gen$Tempo)
```

#4) Remove Entries with NA Values
```{r}
Music_Gen <- na.omit(Music_Gen)
```

#5) Index the ID numbers
```{r}
rownames(Music_Gen)<-Music_Gen$ID
```

#6) Get Rid of Id, Artist, and non numerical values
```{r}
New_Music_Gen <- Music_Gen[,-1:-3]
New_Music_Gen$Instrumentalness <- NULL
New_Music_Gen$Key <- NULL
New_Music_Gen$Genre <- NULL
New_Music_Gen$Mode <- NULL
New_Music_Gen$Acoustics <- NULL
New_Music_Gen$Duration_ms <- NULL
New_Music_Gen$Loudness <- NULL
New_Music_Gen$Lyrics <- NULL
New_Music_Gen$Liveness <- NULL
New_Music_Gen$Obtained_Date <- NULL
```


#7) WSS Plot create function and run
```{r}
wssplot <- function(data, nc=10, seed=1234)
{
  wss <- (nrow(data)-1)*sum(apply(data,2,var))
  for (i in 2:nc){
  set.seed(seed)
  wss[i] <- sum(kmeans(data,centers=i)$withinss)}
  plot(1:nc, wss, type="b", xlab= "Number of Clusters",
  ylab = "Within groups Sum of Squares")
}
wssplot(New_Music_Gen)
```

#8) K-means Cluster / Plot / Centers
```{r}
KM = kmeans(New_Music_Gen,2)
autoplot(KM, New_Music_Gen, frame= TRUE)
KM$centers
```
################################## K-Means Cluster     END    ##############################################









####################### Test And Train [Still Working on This] #########################################

#1) Split Data into Test and Training (Popularity)
```{r}
split = sample.split(New_Music_Gen$Popularity, SplitRatio = .80)
train = subset(New_Music_Gen, split == TRUE)
test = subset(New_Music_Gen, split == FALSE)
```

Music_Pop_dt_model <- train(Popularity ~ ., data = train, metric = "Accuracy", method = "rpart")
Music_Pop_dt_predict <- predict(Music_Pop_dt_model, newdata = test, na.action = na.omit, type = "prob")
head(Music_Pop_dt_predict, 12)

########################### Test And Trai END ##########################################

